Jahn-Teller effect with rigid octahedral rotations in perovskites

### 2023-09-17 manuscript/paper.tex
JT, cooperative...

%\subsection{Examples}
To complete the revision of the JT instability,
let's consider %the following two cases.
%\begin{enumerate}
%\item
a JT/B ion with $2$ electrons.
In the symmetric octahedron case, 
we will have a degenerate (many electron) ground state 
with both electron in the same spin state (Hund's rule) in the $t_{2g}$ manifold.
Distorting the structure $\delta>0$, we lower the ground state energy because 
we can now choose $yz,zx$ orbitals that have lower energy than the original $t_{2g}$ energy of the undistorted structure.
%\item
In case of a JT/B ion with $3$ electrons, the situation is very different.
Assuming Hund's coupling is enough to still produce complete spin polarisation,
their is only one possible orbital configuration, one electron in each of $xy,yz,zx$ orbitals.
This non-degenerate (many electron) ground state
does not lower its energy if the structure is distorted 
because the gain due to $yz,zx$ is already balanced by the loss due to picking the higher energy $xy$ state.
%\end{enumerate}

@@

%\subsection{cooperative JT and orbital order}
\az{just a brief description... cooperative JT and orbital order}

>>>>

%\subsection{Examples}
Let us consider %the following two cases.
%\begin{enumerate}
%\item
a JT/B ion with two electrons.
In the symmetric octahedron case, 
we will have a degenerate (many electron) ground state 
with both electron in the same spin state (Hund's rule) in the $t_{2g}$ manifold.
Distorting the structure $\delta>0$, we lower the ground state energy because 
we can now choose $yz,zx$ orbitals that have lower energy than the original $t_{2g}$ energy of the undistorted structure.
%\item
In case of a JT/B ion with three electrons, the situation is very different.
Assuming Hund's coupling is enough to still produce complete spin polarisation,
their is only one possible orbital configuration, one electron in each of $xy,yz,zx$ orbitals.
This non-degenerate (many electron) ground state
does not lower its energy if the structure is distorted 
because the gain due to $yz,zx$ is already balanced by the loss due to picking the higher energy $xy$ state.
%\end{enumerate}

%\subsection{cooperative JT and orbital order}
%\az{just a brief description... cooperative JT and orbital order}
To complete the revision of the JT instability,
it is worth mentioning that
JT effect occurs in all kinds of systems, molecules, clusters and crystals.
In crystals, e.g., perovskites, it becomes a cooperative effect where the
elongation of one octahedron is accompanied by a compression of adjacent octahedron,
which minimises the changes in the unit cell volume and hence elastic energy cost of the transformation.
Such alternate elongations and compressions also lead to 
an interesting long range orbital ordering where alternate orbitals
are populated on the adjacent JT sites.

@@
